### 2022-01-13 output/manuscript_2022-01-13.tex
literature review

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{TBD} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Influence of School Closure on COVID-19 Contaminations in Ireland and Repercussions across Age Groups},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{Influence of School Closure on COVID-19 Contaminations in
Ireland and Repercussions across Age Groups}
    \author[Dublin City University]{Edgar Morgenroth\corref{1}}
   \ead{edgar.morgenroth@dcu.ie} 
    \author[Dublin City University]{Damien Dupré\corref{2}}
   \ead{damien.dupre@dcu.ie} 
      \address[Dublin City University]{Glasnevin, Dublin 9, Ireland}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

  \begin{abstract}
  TBD
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Since the onset of Covid-19, contradicting messages are circulating
about the spread of the virus in schools. Sometimes presented as a safe
environment (Falk et al., 2021; Walger et al., 2020), at other times
designed as an aggravating factor of the pandemic (Lopez et al., 2020;
Meuris et al., 2021), child care facilities and schools have been closed
multiple times in most countries to protect the public health. However,
the efficacy of school closure is still unknown and little is know about
how Covid-19 Cases spread across age cohort and especially from the
younger cohort being at school to older cohorts including parents, grand
parents and relatives. Some evidence show that school is an influential
environment where the virus is likely to spread. Once contaminated at
school, a common spread believe is that the children and teenagers will
then contaminate their parents. Therefore, among the political actions
to stop the spread of Covid-19, the closure of primary and secondary
schools was widely adopted worldwide. This research aims to investigate
1) the effect of school closure on the evolution of Covid-19 cases and
2) the temporal relationship between the rise of cases in children and
teenagers with the rise of cases of older age group.

\hypertarget{influence-of-school-closure}{%
\subsection{Influence of School
Closure}\label{influence-of-school-closure}}

One of the strongest assumption of government actions is that by
reducing social contact will reduce the spread of the virus. Therefore,
it is legitimate to believe that by closing schools, a reduction of the
contaminations would be observed in the younger age groups. However, the
efficiency of school closure on the reduction of Covid-19 cases is still
questioned. While some research have observed that school closures
contribute to limit or to reduce the growth rate of confirmed cases
after implementation (Stage et al., 2021; Sugishita, 2020), other did
not observe a change in the evolution of Covid-19 cases (Chang et al.,
2020; Iwata et al., 2020). For instance, a controlled comparison between
similar localities in Japan with schools closed and school open did not
revealed any evidence that school closures reduced the spread of
Covid-19 (Fukumoto et al., 2021). If the school closure had a real
impact on the evolution of confirmed Covid-19 cases, it should be
possible to observe a decrease or a least an inflection in the trend of
its evolution among younger age groups.

\hypertarget{causal-relationship-between-age-groups}{%
\subsection{Causal Relationship Between Age
Groups}\label{causal-relationship-between-age-groups}}

A second implicit believe regarding the effect of school closure on the
spread of Covid-19 is that school not only has an effect on to the
spread of the virus in children and teenagers but also has a knock-on
effect on the spread of the virus in older age groups also called
Secondary Attack Rate (SAR). The contaminated children and teenagers
would bring the virus back at home and, then, they will contaminate
their parents and relatives. Despite being reported in several clinical
and epidemiological studies (Siebach et al., 2021; Zhen-Dong et al.,
2020), multiple research have shown that the SAR from children to
household members was, in fact, lower than expected (Heavey et al.,
2020; Hoek et al., 2020; Kim et al., 2021; Ludvigsson, 2020). However,
the SAR of children and teenagers to the household member is likely to
be age-dependent, with difference between infants, primary and secondary
school children, and college students (Gras-Le Guen et al., 2021). A
research investigating the contamination in the household network not
only revealed an exceptional high rate of secondary contamination but
also that these contamination happened when the school were closed
(Soriano-Arandes et al., 2021).

\hypertarget{method}{%
\section{Method}\label{method}}

Contrary to Granger Causality, Transfer Entropy is robust to
non-stationary time series (Silini and Masoller, 2021). The Transfer
Entropy analysis of younger age groups daily changes in Covid-19 cases
on older age groups is done using the R package RTransferEntropy
(Behrendt et al., 2019).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Knowledge about the transmission of the virus significantly improved
with the amount of studies performed, especially in the case of how the
virus behaves with children. From the early analyses showing that the
virus was instantness in children (Li et al., 2020), the results has
changed to more nuanced position which states that the spread of the
virus in children is moderate.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-behrendt2019rtransferentropy}{}%
Behrendt, S., Dimpfl, T., Peter, F.J., Zimmermann, D.J., 2019.
RTransferEntropy --- quantifying information flow between different time
series using effective transfer entropy. SoftwareX 10, 100265.
doi:\href{https://doi.org/10.1038/s41598-021-87818-3}{10.1038/s41598-021-87818-3}

\leavevmode\hypertarget{ref-chang2020modelling}{}%
Chang, S.L., Harding, N., Zachreson, C., Cliff, O.M., Prokopenko, M.,
2020. Modelling transmission and control of the COVID-19 pandemic in
australia. Nature communications 11, 1--13.
doi:\href{https://doi.org/10.1038/s41467-020-19393-6}{10.1038/s41467-020-19393-6}

\leavevmode\hypertarget{ref-falk2021covid}{}%
Falk, A., Benda, A., Falk, P., Steffen, S., Wallace, Z., Høeg, T.B.,
2021. COVID-19 cases and transmission in 17 k--12 schools---wood county,
wisconsin, august 31--november 29, 2020. Morbidity and Mortality Weekly
Report 70, 136--140.
doi:\href{https://doi.org/10.15585/mmwr.mm7004e3}{10.15585/mmwr.mm7004e3}

\leavevmode\hypertarget{ref-fukumoto2021no}{}%
Fukumoto, K., McClean, C.T., Nakagawa, K., 2021. No causal effect of
school closures in japan on the spread of COVID-19 in spring 2020.
Nature medicine 27, 2111--2119.
doi:\href{https://doi.org/10.1038/s41591-021-01571-8}{10.1038/s41591-021-01571-8}

\leavevmode\hypertarget{ref-grasleguen2021reopening}{}%
Gras-Le Guen, C., Cohen, R., Rozenberg, J., Launay, E., Levy-Bruhl, D.,
Delacourt, C., 2021. Reopening schools in the context of increasing
COVID-19 community transmission: The french experience. Archives de
Pédiatrie 28, 178--185.
doi:\href{https://doi.org/10.1016/j.arcped.2021.02.001}{10.1016/j.arcped.2021.02.001}

\leavevmode\hypertarget{ref-heavey2020evidence}{}%
Heavey, L., Casey, G., Kelly, C., Kelly, D., McDarby, G., 2020. No
evidence of secondary transmission of COVID-19 from children attending
school in ireland, 2020. Eurosurveillance 25.
doi:\href{https://doi.org/10.2807/1560-7917.ES.2020.25.21.2000903}{10.2807/1560-7917.ES.2020.25.21.2000903}

\leavevmode\hypertarget{ref-vanderhoek2020role}{}%
Hoek, W. van der, Backer, J.A., Bodewes, R., Friesema, I., Meijer, A.,
Pijnacker, R., Reukers, D.F.M., Reusken, C., Roof, I., Rots, N., Te
Wierik, M.J.M., Gageldonk-Lafeber, A. van, Waegemaekers, C., Hof, S. van
den, 2020. The role of children in the transmission of SARS-CoV-2.
Nederlands tijdschrift voor geneeskunde 164.

\leavevmode\hypertarget{ref-iwata2020school}{}%
Iwata, K., Doi, A., Miyakoshi, C., 2020. Was school closure effective in
mitigating coronavirus disease 2019 (COVID-19)? Time series analysis
using bayesian inference. International Journal of Infectious Diseases
99, 57--61.
doi:\href{https://doi.org/10.1016/j.ijid.2020.07.052}{10.1016/j.ijid.2020.07.052}

\leavevmode\hypertarget{ref-kim2021role}{}%
Kim, J., Choe, Y.J., Lee, J., Park, Y.J., Park, O., Han, M.S., Kim,
J.-H., Choi, E.H., 2021. Role of children in household transmission of
COVID-19. Archives of Disease in Childhood 106, 709--711.
doi:\href{https://doi.org/10.1136/archdischild-2020-319910}{10.1136/archdischild-2020-319910}

\leavevmode\hypertarget{ref-li2020role}{}%
Li, X., Xu, W., Dozier, M., He, Y., Kirolos, A., Theodoratou, E.,
others, 2020. The role of children in transmission of SARS-CoV-2: A
rapid review. Journal of global health 10.
doi:\href{https://doi.org/10.7189/jogh.10.011101}{10.7189/jogh.10.011101}

\leavevmode\hypertarget{ref-lopez2020transmission}{}%
Lopez, A.S., Hill, M., Antezano, J., Vilven, D., Rutner, T., Bogdanow,
L., Claflin, C., Kracalik, I.T., Fields, V.L., Dunn, A., others, 2020.
Transmission dynamics of COVID-19 outbreaks associated with child care
facilities---salt lake city, utah, april--july 2020. Morbidity and
Mortality Weekly Report 69, 1319.
doi:\href{https://doi.org/10.15585/mmwr.mm6937e3}{10.15585/mmwr.mm6937e3}

\leavevmode\hypertarget{ref-ludvigsson2020children}{}%
Ludvigsson, J.F., 2020. Children are unlikely to be the main drivers of
the COVID-19 pandemic -- a systematic review. Acta Paediatrica 109,
1525--1530. doi:\url{https://doi.org/10.1111/apa.15371}

\leavevmode\hypertarget{ref-meuris2021transmission}{}%
Meuris, C., Kremer, C., Geerinck, A., Locquet, M., Bruyère, O., Defêche,
J., Meex, C., Hayette, M.-P., Duchene, L., Dellot, P., Azarzar, S.,
Maréchal, N., Sauvage, A.-S., Frippiat, F., Giot, J.-B., Léonard, P.,
Fombellida, K., Moutschen, M., Durkin, K., Artesi, M., Bours, V., Faes,
C., Hens, N., Darcis, G., 2021. Transmission of SARS-CoV-2 after
COVID-19 screening and mitigation measures for primary school children
attending school in liège, belgium. JAMA Network Open 4,
e2128757--e2128757.
doi:\href{https://doi.org/10.1001/jamanetworkopen.2021.28757}{10.1001/jamanetworkopen.2021.28757}

\leavevmode\hypertarget{ref-siebach2021childhood}{}%
Siebach, M.K., Piedimonte, G., Ley, S.H., 2021. COVID-19 in childhood:
Transmission, clinical presentation, complications and risk factors.
Pediatric Pulmonology 56, 1342--1356.
doi:\url{https://doi.org/10.1002/ppul.25344}

\leavevmode\hypertarget{ref-silini2021fast}{}%
Silini, R., Masoller, C., 2021. Fast and effective pseudo transfer
entropy for bivariate data-driven causal inference. Scientific Reports
11, 1--13.
doi:\href{https://doi.org/10.1016/j.softx.2019.100265}{10.1016/j.softx.2019.100265}

\leavevmode\hypertarget{ref-sorianoarandes2021household}{}%
Soriano-Arandes, A., Gatell, A., Serrano, P., Biosca, M., Campillo, F.,
Capdevila, R., Fàbrega, A., Lobato, Z., López, N., Moreno, A.M., Poblet,
M., Riera-Bosch, M.T., Rius, N., Ruiz, M., Sánchez, A., Valldepérez, C.,
Vilà, M., Pineda, V., Lazcano, U., Díaz, Y., Reyes-Urueña, J.,
Soler-Palacín, P., Catalonia Research Group, C.P.D. in, 2021. Household
severe acute respiratory syndrome coronavirus 2 transmission and
children: A network prospective study. Clinical Infectious Diseases 73,
e1261--e1269.
doi:\href{https://doi.org/10.1093/cid/ciab228}{10.1093/cid/ciab228}

\leavevmode\hypertarget{ref-stage2021shut}{}%
Stage, H.B., Shingleton, J., Ghosh, S., Scarabel, F., Pellis, L.,
Finnie, T., 2021. Shut and re-open: The role of schools in the spread of
COVID-19 in europe. Philosophical Transactions of the Royal Society B
376, 20200277.
doi:\href{https://doi.org/10.1098/rstb.2020.0277}{10.1098/rstb.2020.0277}

\leavevmode\hypertarget{ref-yoshiyuki2020effects}{}%
Sugishita, J.A.S., Yoshiyuki AND Kurita, 2020. Effects of voluntary
event cancellation and school closure as countermeasures against
COVID-19 outbreak in japan. PLOS ONE 15, 1--10.
doi:\href{https://doi.org/10.1371/journal.pone.0239455}{10.1371/journal.pone.0239455}

\leavevmode\hypertarget{ref-walger2020children}{}%
Walger, P., Heininger, U., Knuf, M., Exner, M., Popp, W., Fischbach, T.,
Trapp, S., Hübner, J., Herr, C., Simon, A., others, 2020. Children and
adolescents in the CoVid-19 pandemic: Schools and daycare centers are to
be opened again without restrictions. The protection of teachers,
educators, carers and parents and the general hygiene rules do not
conflict with this. GMS hygiene and infection control 15.
doi:\href{https://doi.org/10.3205/dgkh000346}{10.3205/dgkh000346}

\leavevmode\hypertarget{ref-zhendong2020clinical}{}%
Zhen-Dong, Y., Gao-Jun, Z., Run-Ming, J., Zhi-Sheng, L., Zong-Qi, D.,
Xiong, X., Guo-Wei, S., 2020. Clinical and transmission dynamics
characteristics of 406 children with coronavirus disease 2019 in china:
A review. Journal of Infection 81, e11--e15.
doi:\href{https://doi.org/10.1016/j.jinf.2020.04.030}{10.1016/j.jinf.2020.04.030}

\end{CSLReferences}

\end{document}