### 2022-02-10 manuscript/appendix_1.Rmd
europe

# Augmented Dickey-Fuller test for stationay timeseries

```{r}
adf_res <- df_long |>
  group_by(age_group) |>
  summarise(
    aTSA::adf.test(covid_changes, output = FALSE) |>
      map_df(as_tibble)
  )
```