### 2023-12-18 manuscript_elsevier/appendix_1.Rmd
manuscript elsevier

---
title: "Appendix 1: Dickey-Fuller test for stationay timeseries"
toc: false

output:
  bookdown::pdf_book

header-includes:
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{threeparttable}
  - \floatplacement{figure}{H}
  - \floatplacement{table}{H}
  - \usepackage{caption}
---

Test for the null hypothesis of a unit root of a univarate time series (equivalently, non-stationary time series).

The Augmented Dickey-Fuller test incorporates three types of linear regression models. The first type is a linear model with no drift and linear trend with respect to time. The second type is a linear model with drift but no linear trend. The third type is a linear model with both drift and linear trend.

The p.value is calculated by interpolating the test statistics from the corresponding critical values tables (Fuller, 1996) for each type of linear models with given sample size.

```{r setup, include = FALSE}
# libraries --------------------------------------------------------------------
library(here)
library(tidyverse)
library(knitr)
library(kableExtra)

# options ----------------------------------------------------------------------
set.seed(123) # Seed for random number generation
options(scipen = 999) # disable sci number format
opts_chunk$set(
  cache.extra = rand_seed, 
  cache = FALSE,
  message = FALSE, 
  warning = FALSE, 
  error = FALSE,
  echo = FALSE, 
  fig.retina = 3,
  dev = "ragg_png"
  )
```

```{r}
df_global_age <- read_csv(here("data/df_global_age.csv"))

adf_res_type1 <- df_global_age |>
  group_by(age_grp, country) |>
  group_modify(~ aTSA::adf.test(.x$daily_changes, output = FALSE, nlag = 2) |>
                 bind_rows() |> 
                 pluck("type1") |> 
                 as_tibble()
  ) |> 
  rename(ADF_type1 = ADF, p.value_type1 = p.value)

adf_res_type2 <- df_global_age |>
  group_by(age_grp, country) |>
  group_modify(~ aTSA::adf.test(.x$daily_changes, output = FALSE, nlag = 2) |>
                 bind_rows() |> 
                 pluck("type2") |> 
                 as_tibble()
  ) |> 
  rename(ADF_type2 = ADF, p.value_type2 = p.value)

adf_res_type3 <- df_global_age |>
  group_by(age_grp, country) |>
  group_modify(~ aTSA::adf.test(.x$daily_changes, output = FALSE, nlag = 2) |>
                 bind_rows() |> 
                 pluck("type3") |> 
                 as_tibble()
  ) |> 
  rename(ADF_type3 = ADF, p.value_type3 = p.value)

adf_res <- adf_res_type1 |> 
  full_join(adf_res_type2) |> 
  full_join(adf_res_type3)
```

```{r}
adf_res |>
  filter(age_grp %in% c("aged0to4", "aged5to9")) |> 
  kbl(
    caption = "Dickey-Fuller test for all combinaisons of countries and age groups investigated",
    col.names = c("Age Group", "Country", "Lag", "ADF", "p", "ADF", "p", "ADF", "p")
  ) |> 
  kable_classic(full_width = F) |> 
  add_header_above(c(" " = 3, "Type 1" = 2, "Type 2" = 2, "type 3" = 2)) |>
  collapse_rows(columns = 1:2, valign = "middle")
```

\newpage

```{r}
adf_res |>
  filter(age_grp %in% c("aged10to14", "aged15to19")) |> 
  kbl(
    caption = "Dickey-Fuller test for all combinaisons of countries and age groups investigated",
    col.names = c("Age Group", "Country", "Lag", "ADF", "p", "ADF", "p", "ADF", "p")
  ) |> 
  kable_classic(full_width = F) |> 
  add_header_above(c(" " = 3, "Type 1" = 2, "Type 2" = 2, "type 3" = 2)) |>
  collapse_rows(columns = 1:2, valign = "middle")
```

\newpage

```{r}
adf_res |>
  filter(age_grp %in% c("aged20to24", "aged25to34")) |> 
  kbl(
    caption = "Dickey-Fuller test for all combinaisons of countries and age groups investigated",
    col.names = c("Age Group", "Country", "Lag", "ADF", "p", "ADF", "p", "ADF", "p")
  ) |> 
  kable_classic(full_width = F) |> 
  add_header_above(c(" " = 3, "Type 1" = 2, "Type 2" = 2, "type 3" = 2)) |>
  collapse_rows(columns = 1:2, valign = "middle")
```

\newpage

```{r}
adf_res |>
  filter(age_grp %in% c("aged35to44", "aged45to54")) |> 
  kbl(
    caption = "Dickey-Fuller test for all combinaisons of countries and age groups investigated",
    col.names = c("Age Group", "Country", "Lag", "ADF", "p", "ADF", "p", "ADF", "p")
  ) |> 
  kable_classic(full_width = F) |> 
  add_header_above(c(" " = 3, "Type 1" = 2, "Type 2" = 2, "type 3" = 2)) |>
  collapse_rows(columns = 1:2, valign = "middle")
```

\newpage

```{r}
adf_res |>
  filter(age_grp %in% c("aged55to64", "aged65+")) |> 
  kbl(
    caption = "Dickey-Fuller test for all combinaisons of countries and age groups investigated",
    col.names = c("Age Group", "Country", "Lag", "ADF", "p", "ADF", "p", "ADF", "p")
  ) |> 
  kable_classic(full_width = F) |> 
  add_header_above(c(" " = 3, "Type 1" = 2, "Type 2" = 2, "type 3" = 2)) |>
  collapse_rows(columns = 1:2, valign = "middle")
```