### 2025-06-05 manuscript_njd/template.qmd

---
title: "Quarto Template (wiley-njd) Example"

format:
  wiley-njd-pdf:
    article-type: "Article Type: Exemplar"
    received: "10 November 2024"
    revised: "10 November 2024"
    accepted: "1 January 2025"
    journal: "Journal: None"
    volume: "00"
    copy-year: 2024
    start-page: 1024

    titlemark: "Short Title (in header)"
    authormark: "Short Author (in header)"

    keywords:
      - keyword1
      - buzzword
      - keyword3

author:
  - name: Josh Cowley
    email: joshcowley@fakeemails.com
    corresponding: true
    affiliations:
      - name: That University
        department: School of Something
        city: City
        country: UK

  - name: The Supervisor
    email: thesupervisor@fakeemails.com
    affiliations:
      - name: That University
        department: School of Something
        city: City
        country: UK

  - name: Business Contact
    email: businesscontact@fakebusinessemails.com
    affiliations:
      - name: That Company
        department: Advanced Department
        city: City
        country: Germany

present-address: |
  The present address is currently hidden.

abstract: |
  Document serves as example input for Quarto template openly available
    from Github, see more information at
    <https://github.com/nclJoshCowley/wiley-njd>.
  All rendering requires LaTeX to be installed, this includes any 
    TeX installation bundled with Quarto.
  Since Wiley's "New Journal Design" (NJD) LaTeX template is utilised in
    the build process, usage is subject to any license associated with
    original template.

keep-tex: true
bibliography: bibliography.bib
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 10))

library(knitr)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  # Example journal specifications: (width: 17.15 cm, DPI: >= 300)
  fig.width = 17.15 / 2.54,
  dpi = 300
)
```

# Example Content

## Raw Latex {#sec-rawlatex}

Any required LaTeX can be added to the document as required.
For Quarto, this is best done by including the raw LaTeX inside a 
  ` ```{=latex} ... ``` ` block, as shown in the source document.

```{=latex}
\setlength{\parindent}{0pt}
```

## Equations {#sec-eqns}

As with LaTeX, we can create in-line equations with $y = mx + c$ and create
  display single-line equations with

$$
    a^n
        = b^n + c^n
$$

For multi-line equations, use the `$$` around a `begin{aligned}[t]` method.

$$
\begin{aligned}[t]
    y_1
        &= \beta_0 + \beta_1 x_1
    \\
        &\vdots
    \\
    y_n
        &= \beta_0 + \beta_1 x_n
\end{aligned}
$$ {#eq-linreg}

## Code

In-line code can be enacted with backticks, for example $1 + 1 = `r 1 + 1`$.
Display code chunks are formulated in the usual Quarto way.

```{r}
factorial(5)
```

For example, output is suppressed via the chunk option `#| echo: false`.

```{r}
#| echo: false
sprintf("10! = %i", factorial(10))
```

## Images

We can include graphics that already exist with `knitr::include_graphics`
  or by `\includegraphics` using the method described in @sec-rawlatex.

```{r}
knitr::include_graphics("https://quarto.org/quarto.png")
```

And we can generate our own plots using R code.
Note that we only change `fig.width` and `fig.height` or `fig.asp`;
  changing `out.height` or `out.width` will stretch the image.

```{r}
#| label: fig-cars
#| fig-cap: "A scatter plot of weight and MPG from the `mtcars` dataset"
#| fig-asp: 0.8
#| echo: false
mtcars |> 
  ggplot2::ggplot(ggplot2::aes(x = mpg, y = wt)) +
  ggplot2::geom_smooth(method = "lm", formula = y ~ x, colour = "red") +
  ggplot2::geom_point() +
  ggplot2::labs(
    title = "Negative Correlation between Weight and MPG",
    y = "Weight",
    x = "Miles per Gallon (MPG)"
  )
```

## Cross-Reference

We want to be able to reference already defined items such as

1. linear regression, as described in (@eq-linreg) from @sec-eqns.

1. scatter plot of `mtcars` data, @fig-cars.

## Citations

We can cite is a few different ways that are consistent with `natbib`

1. Parenthesised citation, using `citep` [@Hirt1974].

1. Textual citation, using `citet` as in @Hirt1974.

Multiple citations follow (semi-colon separated) Quarto syntax
  [@Hirt1974; @Benson1992; @Dukowicz1984].

## Footnotes

Here is a footnote reference,[^1] and another.[^longnote]

[^1]: Here is the footnote.

[^longnote]: Here's one with multiple blocks.

    Subsequent paragraphs are indented to show that they belong to the previous
    footnote.

    Multi-paragraph footnotes work like multi-paragraph list items.

## Quotes

The `quote` environment is supplied by the Markdown quote syntax where lines
  begin with `>`.
For example,

> The combination of some data and an aching desire for an answer does not
> ensure that a reasonable answer can be extracted from a given body of data.
> 
> -- John Tukey

\newpage

# Future Work

## NJD Consistency

Wiley's NJD supplies

* 2 column layouts, only 1 column layouts apply here.

* Multiple referencing systems, only Harvard applies here.

* Font options (Helvetica, Lato, etc.) but installing fonts is non-trivial.
  Currently Times is used but we may want sans-serif.

## Theorems

Custom environments supplied by NJD should work with Quarto divs, however
  further testing is required.

::: {#thm-1}
This is a theorem.
:::

::: {proof}
With a proof.
:::

::: {#prp-1}
This is a proposition.
:::

::: {#def-1}
This is a definition
:::

\newpage

# References

::: {#refs}
:::